---
title: "2023_08_11_Final"
output: html_document
date: "2023-08-11"
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=TRUE, results = "html")
library(psych)
#library(lme4)
library(ggplot2)
library(sjPlot)
library(ggeffects)
library(arrow)
library(ggpubr)
library(tidyverse)
#library(miceadds)
library(sandwich)
library(lmtest)
library(emmeans)
library(dplyr)
```

```{r load_data ,warning=FALSE, echo = FALSE}
data_root <- "../../data"
data <- read_parquet(file.path(data_root, "cleaned_data", "2023-06-18_long_form.parquet"))
```

```{r recode, echo = FALSE}
# Recode for regression
data$exp_response <-ifelse(data$exp_response == TRUE, 1, 0)
data$veracity<-as.factor(data$veracity)
data$Condition<-as.factor(data$Condition)
data$option_cond <-as.factor(data$option_cond)
data$congruency <-as.factor(data$congruency)
data$congruency <-ifelse(data$congruency == "incongruent", 1, 0)

data$age <- 2023 - data$year_birth
data$age_recode <- data$age/10

# Captures the five headline scenarios observed in our data
data<- data %>% mutate(
  cGPT_accuracy_recode = case_when(
    veracity == TRUE & ano_true_false_unsure == "false" ~ "True_false",
    veracity == FALSE & ano_true_false_unsure == "false" ~ "False_false",
    veracity == TRUE & ano_true_false_unsure == "unsure" ~ "True_unsure",
    veracity == FALSE & ano_true_false_unsure == "unsure" ~ "False_unsure",
    veracity == TRUE & ano_true_false_unsure == "true" ~ "True_true",
  ))

# data$cGPT_accuracy_recode <-as.factor(data$cGPT_accuracy_recode)

data = data[,c("exp_response", "Group", "veracity", "Condition", "option_cond", "congruency", "cGPT_accuracy_recode", "age", "edu_recoded", "ResponseId", "qualtrics_question_num", "AI_att_mean")]

# Split belief and share datasets
belief_data <- subset(data, data$Group == "Belief")
share_data <- subset(data, data$Group == "Share")

# Create version WITHOUT optional participants
belief_data_no_opt <- belief_data[belief_data$Condition != "Optional",]
share_data_no_opt <- share_data[share_data$Condition != "Optional",]

# Create version with ONLY optional participants
belief_data_opt<- belief_data[belief_data$Condition == "Optional",]
share_data_opt<- share_data[share_data$Condition == "Optional",]
```
### Note: 
- 1) The post hoc estimates are printed before conducting pairwise comparison
- 2) Clustered standard errors were included for pairwise comparison. 


### Contents:
- [A. Confirm main text findings]
	- [Set equations (confirm)]
	- [A.1a Confirm main effects: Belief]
	- [A.1b Confirm main effects: Share]
	- [A.2a Confirm results accounting for AI judgement: Belief]
	- [A.2b Confirm results accounting for AI judgement: Share]
	- [A.3a Confirm results for Optional condition: Belief]
	- [A.3b Confirm results for Optional condition: Share]
- [B. Interactions]
	- [Congruency]
		- [Set equations (congruency)]
		- [B.1a Main effects w. congruency: Belief]
		- [B.1b Main effects w. congruency: Share]
		- [B.2a Accounting for AI judgement w. congruency: Belief]
		- [B.2b Accounting for AI judgement w. congruency: Share]
		- [B.3a Optional w. congruency: Belief]
		- [B.3b Optional w. congruency: Share]
	- [Attitude towards AI (ATAI)]
		- [Set equations (ATAI)]
		- [B.4a Main effects w. ATAI: Belief]
		- [B.4b Main effects w. ATAI: Share]
		- [B.5a Accounting for AI judgement w. ATAI: Belief]
		- [B.6b Accounting for AI judgement w. ATAI: Share]
		- [B.7a Optional w. ATAI: Belief]
		- [B.7b Optional w. ATAI: Share]

### Set all pairs of interest utilized throughout the code
```{r POI}
ai_judge_pairs <-c("Control False_false - Forced False_false", "Control False_unsure - Forced False_unsure", "Control True_false - Forced True_false", "Control True_true - Forced True_true", "Control True_unsure - Forced True_unsure" )
optional_pairs <-c( "Opt_in FALSE - Opt_out FALSE", "Opt_in TRUE - Opt_out TRUE")
```

### Define function to print model stats
```{r p_func}
# Define function to extract overall p-value of model
model_p <- function(model_summary) {
    f <- model_summary$fstatistic
	p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL # Ensures the value is numeric
    if (p == 0) {
        p <- "< 2.2e-16"
    }
    return(p)
}

print_summary <- function(model) {
	model_summary <- summary(model)
	f_statistic <- model_summary$fstatistic[1]
	cat("F statistic:", f_statistic, "\n")
	
	# df_model <- model_summary$df[1]
	# cat("Degrees of freedom (model):", df_model, "\n")
	
	df_residual <- model_summary$df[2]
	cat("Degrees of freedom (residuals):", df_residual, "\n")
	
	p_value <- model_p(model_summary)
	cat("P-value:", p_value, "\n")
	
	r2 <- model_summary$r.squared
	cat("R-squared:", r2, "\n")
	
	r2adj <- model_summary$adj.r.squared
	cat("Adjusted R-squared:", r2adj, "\n")
}

```

# A. Confirm main text findings

### Set equations (confirm)
```{r set_equations}
main_effects_formula = exp_response ~ Condition*veracity + age + edu_recoded
five_way_formula = exp_response ~ Condition*cGPT_accuracy_recode + age + edu_recoded
optional_formula = exp_response ~ option_cond * veracity + age + edu_recoded
```

### A.1a Confirm main effects: Belief
```{r confirm_me_belief}
mb_me <- lm(main_effects_formula, data = belief_data)

print_summary(mb_me)

clustered_se <- vcovCL(mb_me, cluster = belief_data[, c("ResponseId", "qualtrics_question_num")])
coeftest(mb_me, vcov = clustered_se)
```

### A.1b Confirm main effects: Share

```{r confirm_me_share}
ms_me <- lm(main_effects_formula, data = share_data)

print_summary(ms_me)

clustered_se <- vcovCL(ms_me, cluster = share_data[, c("ResponseId", "qualtrics_question_num")])
coeftest(ms_me, vcov = clustered_se)
```

### A.2a Confirm results accounting for AI judgement: Belief

#### Fit regression model
```{r confirm_w_AI_judgement_belief}
mb_ai <- lm(five_way_formula, data = belief_data_no_opt)

print_summary(mb_ai)

clustered_se <- vcovCL(mb_ai, cluster = belief_data_no_opt[, c("ResponseId", "qualtrics_question_num")])
coeftest(mb_ai, vcov = clustered_se)
```



#### Run post-hoc tests

```{r AI_judgement_post_hoc_belief}
test<-emmeans(mb_ai, ~ Condition* cGPT_accuracy_recode, infer = TRUE)
test

pairwise <- contrast(test, method = "pairwise", infer = c(TRUE, TRUE), adjust = "bonferroni", vcov = clustered_se)
ph_belief <- as.data.frame(pairwise) 
res <- ph_belief[ph_belief$contrast %in% ai_judge_pairs,]
res$p.value <- round(res$p.value, 5)
res
```


### A.2b Confirm results accounting for AI judgement: Share

#### Fit regression model
```{r confirm_w_AI_judgement_share}
ms_ai <- lm(five_way_formula, data = share_data_no_opt)

print_summary(ms_ai)

clustered_se <- vcovCL(ms_ai, cluster = share_data_no_opt[, c("ResponseId", "qualtrics_question_num")])
coeftest(ms_ai, vcov = clustered_se, cluster = share_data_no_opt[, c("ResponseId", "qualtrics_question_num")])
```

#### Run post-hoc tests

```{r AI_judgement_post_hoc_share}
test<-emmeans(ms_ai, ~ Condition* cGPT_accuracy_recode, infer = TRUE)
test
pairwise <- contrast(test, method = "pairwise", infer = c(TRUE, TRUE), adjust = "bonferroni", vcov = clustered_se)

ph_share <- as.data.frame(pairwise) #pairwise# pairs(test)
res <- ph_share[ph_share$contrast %in% ai_judge_pairs,]
res$p.value <- round(res$p.value, 5)
res

```


### A.3a Confirm results for Optional condition: Belief

#### Fit regression model
```{r confirm_optional_belief}
mb_opt <- lm(optional_formula, data = belief_data_opt)

print_summary(mb_opt)

clustered_se <- vcovCL(mb_opt, cluster = belief_data_opt[, c("ResponseId", "qualtrics_question_num")]) 
coeftest(mb_opt, vcov = clustered_se)
```

#### Run post-hoc tests

```{r optional_post_hoc_belief}
test<- emmeans(mb_opt, ~ option_cond* veracity, infer = TRUE)
test

pairwise <- contrast(test, method = "pairwise", infer = c(TRUE, TRUE), adjust = "bonferroni", vcov = clustered_se)
ph_belief <- as.data.frame(pairwise)
res <- ph_belief[ph_belief$contrast %in% optional_pairs,]
res$p.value <- round(res$p.value, 5)
res
```


### A.3b Confirm results for Optional condition: Share

#### Fit regression model
```{r confirm_optional_share}
ms_opt <- lm(optional_formula, data = share_data_opt )

print_summary(ms_opt)

clustered_se <- vcovCL(ms_opt, cluster = share_data_opt[, c("ResponseId", "qualtrics_question_num")]) 
coeftest(ms_opt, vcov = clustered_se)
```

#### Run post-hoc tests

```{r optional_post_hoc_share}
test<-emmeans(ms_opt, ~ option_cond* veracity, infer= TRUE )
test

pairwise <- contrast(test, method = "pairwise", infer = c(TRUE, TRUE), adjust = "bonferroni", vcov = clustered_se)
ph_share <- as.data.frame(pairwise)
res <- ph_share[ph_share$contrast %in% optional_pairs,]
res$p.value <- round(res$p.value, 5)
res
```

---
---
---

# B. Interactions

## Congruency

### Set equations (congruency)
```{r set_equations_int_congruency}
me_w_congruency_formula = exp_response ~ Condition*veracity*congruency + age + edu_recoded
five_way_congruency_formula = exp_response ~ Condition*cGPT_accuracy_recode*congruency + age + edu_recoded
optional_congruency_formula = exp_response ~ option_cond*veracity*congruency + age + edu_recoded
```

### B.1a Main effects w. congruency: Belief
```{r me_congru_int_belief}
mb_me_congru <- lm(me_w_congruency_formula, data = belief_data)

print_summary(mb_me_congru)

clustered_se <- vcovCL(mb_me_congru, cluster = belief_data[, c("ResponseId", "qualtrics_question_num")])
coeftest(mb_me_congru, vcov = clustered_se)
```

### B.1b Main effects w. congruency: Share

```{r me_congru_int_share}
ms_me_congru <- lm(me_w_congruency_formula, data = share_data)

print_summary(ms_me_congru)

clustered_se <- vcovCL(ms_me_congru, cluster = share_data[, c("ResponseId", "qualtrics_question_num")])
coeftest(ms_me_congru, vcov = clustered_se)
```

### B.2a Accounting for AI judgement w. congruency: Belief

#### Fit regression model
```{r AI_judgement_congruency_belief}
mb_ai_congru <- lm(five_way_congruency_formula, data = belief_data_no_opt)

print_summary(mb_ai_congru)

clustered_se <- vcovCL(mb_ai_congru, cluster = belief_data_no_opt[, c("ResponseId", "qualtrics_question_num")])
coeftest(mb_ai_congru, vcov = clustered_se)
```

#### Post-hoc test
```{r AI_judgement_congruency_post_belief}
test<- emtrends(mb_ai_congru, ~ Condition*cGPT_accuracy_recode, var="congruency", infer = TRUE)
test

pairwise <- contrast(test, method = "pairwise", infer = c(TRUE, TRUE), adjust = "bonferroni", vcov = clustered_se)
ph_belief <- as.data.frame(pairwise)
res <- ph_belief[ph_belief$contrast %in% ai_judge_pairs,]
res$p.value <- round(res$p.value, 5)
res
```


### B.2b Accounting for AI judgement w. congruency: Share

#### Fit regression model
```{r AI_judgement_congruency_share}
ms_ai_congru <- lm(five_way_congruency_formula, data = share_data_no_opt)

print_summary(ms_ai_congru)

clustered_se <- vcovCL(ms_ai_congru, cluster = share_data_no_opt[, c("ResponseId", "qualtrics_question_num")])
coeftest(ms_ai_congru, vcov = clustered_se)
```

#### Post-hoc test
```{r AI_judgement_congruency_post_share}
test<-emtrends(ms_ai_congru, ~ Condition*cGPT_accuracy_recode, var="congruency", infer =TRUE)
test

pairwise <- contrast(test, method = "pairwise", infer = c(TRUE, TRUE), adjust = "bonferroni", vcov = clustered_se)
ph_share <- as.data.frame(pairwise)
res <- ph_share[ph_share$contrast %in% ai_judge_pairs,]
res$p.value <- round(res$p.value, 5)
res
```


### B.3a Optional w. congruency: Belief

#### Fit regression model
```{r optional_congruency_belief}
mb_opt_congru <- lm(optional_congruency_formula, data = belief_data_opt)

print_summary(mb_opt_congru)

clustered_se <- vcovCL(mb_opt_congru, cluster = belief_data_opt[, c("ResponseId", "qualtrics_question_num")]) 
coeftest(mb_opt_congru, vcov = clustered_se)
plot_model(mb_opt_congru, type  = "pred", terms =  c( "congruency", "option_cond",  "veracity"), title = "Moderting effects of headline congurency") + ylab("Believed") + theme_minimal()

```

#### Post-hoc test
```{r optional_congruency_post_belief}
test<-emtrends(mb_opt_congru, ~ option_cond*veracity, var="congruency", infer= TRUE)
test

pairwise <- contrast(test, method = "pairwise", infer = c(TRUE, TRUE), adjust = "bonferroni", vcov = clustered_se)
ph_belief <- as.data.frame(pairwise)
res <- ph_belief[ph_belief$contrast %in% optional_pairs,]
res$p.value <- round(res$p.value, 5)
res
```


### B.3b Optional w. congruency: Share

#### Fit regression model
```{r optional_congruency_share}
ms_opt_congru <- lm(optional_congruency_formula, data = share_data_opt )

print_summary(ms_opt_congru)

clustered_se <- vcovCL(ms_opt_congru, cluster = share_data_opt[, c("ResponseId", "qualtrics_question_num")]) 
coeftest(ms_opt_congru, vcov = clustered_se)
plot_model(ms_opt_congru, type  = "pred", terms =  c( "congruency", "option_cond",  "veracity"), title = "Moderting effects of headline congurency") + ylab("Intention to share") + theme_minimal()

```

#### Post-hoc test
```{r optional_congruency_post_share}
test<-emtrends(ms_opt_congru, ~ option_cond*veracity, var="congruency", infer = TRUE)
test
pairwise <- contrast(test, method = "pairwise", infer = c(TRUE, TRUE), adjust = "bonferroni", vcov = clustered_se)

ph_share <- as.data.frame(pairwise)
res <- ph_share[ph_share$contrast %in% optional_pairs,]
res$p.value <- round(res$p.value, 5)
res
```

---


## Attitude towards AI (ATAI)

### Set equations (ATAI)
```{r set_equations_int_atai}
me_w_atai_formula = exp_response ~ Condition*veracity*AI_att_mean + age + edu_recoded
five_way_atai_formula = exp_response ~ Condition*cGPT_accuracy_recode*AI_att_mean + age + edu_recoded
optional_atai_formula = exp_response ~ option_cond*veracity*AI_att_mean + age + edu_recoded
```

### B.4a Main effects w. ATAI: Belief
```{r me_atai_int_belief}
mb_me_atai <- lm(me_w_atai_formula, data = belief_data)

print_summary(mb_me_atai)

clustered_se <- vcovCL(mb_me_atai, cluster = belief_data[, c("ResponseId", "qualtrics_question_num")])
coeftest(mb_me_atai, vcov = clustered_se)
```

### B.4b Main effects w. ATAI: Share

```{r me_atai_int_share}
ms_me_atai <- lm(me_w_atai_formula, data = share_data)

print_summary(ms_me_atai)

clustered_se <- vcovCL(ms_me_atai, cluster = share_data[, c("ResponseId", "qualtrics_question_num")])
coeftest(ms_me_atai, vcov = clustered_se)
```

### B.5a Accounting for AI judgement w. ATAI: Belief

#### Fit regression model
```{r AI_judgement_atai_belief}
mb_ai_atai <- lm(five_way_atai_formula, data = belief_data_no_opt)

print_summary(mb_ai_atai)

clustered_se <- vcovCL(mb_ai_atai, cluster = belief_data_no_opt[, c("ResponseId", "qualtrics_question_num")])
coeftest(mb_ai_atai, vcov = clustered_se)
```

#### Post-hoc test
```{r AI_judgement_atai_post_belief}
test<-emtrends(mb_ai_atai, ~ Condition*cGPT_accuracy_recode, var="AI_att_mean", infer =TRUE)
test
pairwise <- contrast(test, method = "pairwise", infer = c(TRUE, TRUE), adjust = "bonferroni", vcov = clustered_se)

ph_belief <- as.data.frame(pairwise)
res <- ph_belief[ph_belief$contrast %in% ai_judge_pairs,]
res$p.value <- round(res$p.value, 5)
res
```

### B.6b Accounting for AI judgement w. ATAI: Share

#### Fit regression model
```{r AI_judgement_atai_share}
ms_ai_atai <- lm(five_way_atai_formula, data = share_data_no_opt)

print_summary(ms_ai_atai)

clustered_se <- vcovCL(ms_ai_atai, cluster = share_data_no_opt[, c("ResponseId", "qualtrics_question_num")])
coeftest(ms_ai_atai, vcov = clustered_se)
```

#### Post-hoc test
```{r AI_judgement_atai_post_share}
test<- emtrends(ms_ai_atai, ~ Condition*cGPT_accuracy_recode, var="AI_att_mean", infer = TRUE)
test
pairwise <- contrast(test, method = "pairwise", infer = c(TRUE, TRUE), adjust = "bonferroni", vcov = clustered_se)

ph_share <- as.data.frame(pairwise)
res <- ph_share[ph_share$contrast %in% ai_judge_pairs,]
res$p.value <- round(res$p.value, 5)
res
```

### B.7a Optional w. ATAI: Belief

#### Fit regression model
```{r optional_atai_belief}
mb_opt_atai <- lm(optional_atai_formula, data = belief_data_opt)

print_summary(mb_opt_atai)

clustered_se <- vcovCL(mb_opt_atai, cluster = belief_data_opt[, c("ResponseId", "qualtrics_question_num")]) 
coeftest(mb_opt_atai, vcov = clustered_se)
plot_model(mb_opt_atai, type  = "pred", terms =  c( "AI_att_mean", "option_cond",  "veracity"), title = "Moderting effects of ATAI") + ylab("Believed") + theme_minimal()
```

#### Post-hoc test
```{r optional_atai_post_belief}
test<- emtrends(mb_opt_atai, ~ option_cond*veracity, var="AI_att_mean", infer = TRUE)
test
pairwise <- contrast(test, method = "pairwise", infer = c(TRUE, TRUE), adjust = "bonferroni", vcov = clustered_se)

ph_belief <- as.data.frame(pairwise)
res <- ph_belief[ph_belief$contrast %in% optional_pairs,]
res$p.value <- round(res$p.value, 5)
res
```

### B.7b Optional w. ATAI: Share

#### Fit regression model
```{r optional_atai_share}
ms_ai_atai <- lm(optional_atai_formula, data = share_data_opt )

print_summary(ms_ai_atai)

clustered_se <- vcovCL(ms_ai_atai, cluster = share_data_opt[, c("ResponseId", "qualtrics_question_num")]) 
coeftest(ms_ai_atai, vcov = clustered_se)
plot_model(ms_ai_atai, type  = "pred", terms =  c( "AI_att_mean", "option_cond",  "veracity"), title = "Moderting effects of ATAI") + ylab("Intention to share") + theme_minimal()

```

#### Post-hoc test
```{r optional_atai_post_share}
test<-emtrends(ms_ai_atai, ~ option_cond*veracity, var="AI_att_mean", infer = TRUE)
test
pairwise <- contrast(test, method = "pairwise", infer = c(TRUE, TRUE), adjust = "bonferroni", vcov = clustered_se)

ph_share <- as.data.frame(pairwise)
res <- ph_share[ph_share$contrast %in% optional_pairs,]
res$p.value <- round(res$p.value, 5)
res
```

