---
title: "2024_07_25_interactions_congruency"
date: "2024_07_25"
output: 
  html_document:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=TRUE, results = "html")
library(psych)
#library(lme4)
library(ggplot2)
library(sjPlot)
library(ggeffects)
library(arrow)
library(ggpubr)
library(tidyverse)
#library(miceadds)
library(sandwich)
library(lmtest)
library(emmeans)
library(dplyr)
```

```{r load_data ,warning=FALSE, echo = FALSE}
data_root <- "../data"
data <- read_parquet(file.path(data_root, "cleaned_data", "2024-07-01_long_form_cgpt_fc_paper.parquet"))
```

```{r recode, echo = FALSE}
# Recode for regression
data$exp_response <-ifelse(data$exp_response == TRUE, 1, 0)
data$veracity<-as.factor(data$veracity)
data$Condition<-as.factor(data$Condition)
data$option_cond <-as.factor(data$option_cond)
data$congruency <-as.factor(data$congruency)
data$ano_true_false_unsure <-as.factor(data$ano_true_false_unsure)
data$congruency <-ifelse(data$congruency == "incongruent", 1, 0)

data$age <- ifelse(data$Condition == "Human-FC", 2024 - data$year_birth, 2023 - data$year_birth)
data$age_recode <- data$age/10

# Captures the five headline scenarios observed in our data
data<- data %>% mutate(
  cGPT_accuracy_recode = case_when(
    veracity == TRUE & ano_true_false_unsure == "false" ~ "True_false",
    veracity == FALSE & ano_true_false_unsure == "false" ~ "False_false",
    veracity == TRUE & ano_true_false_unsure == "unsure" ~ "True_unsure",
    veracity == FALSE & ano_true_false_unsure == "unsure" ~ "False_unsure",
    veracity == TRUE & ano_true_false_unsure == "true" ~ "True_true",
  ))

# data$cGPT_accuracy_recode <-as.factor(data$cGPT_accuracy_recode)

data = data[,c("exp_response", "Group", "veracity", "Condition", "option_cond", "congruency", "cGPT_accuracy_recode", "age", "edu_recoded", "ResponseId", "qualtrics_question_num", "AI_att_mean", "ano_true_false_unsure")]

# Split belief and share datasets
belief_data <- subset(data, data$Group == "Belief")
share_data <- subset(data, data$Group == "Share")

# Create version WITHOUT optional or human FC participants
belief_data_no_opt_or_hfc <- belief_data[belief_data$Condition != "Optional" & belief_data$Condition != "Human-FC",]
share_data_no_opt_or_hfc <- share_data[share_data$Condition != "Optional" & share_data$Condition != "Human-FC",]

# Create version with ONLY optional participants
belief_data_opt<- belief_data[belief_data$Condition == "Optional",]
share_data_opt<- share_data[share_data$Condition == "Optional",]
```
## Note: 
- 1) The post hoc estimates are printed before conducting pairwise comparison
- 2) Clustered standard errors were included for pairwise comparison. 


## Contents:
- [Interactions]
	- [Congruency]
		- [Set equations (congruency)]
		- [Main effects w. congruency: Belief]
		- [Main effects w. congruency: Share]
		- [Accounting for AI judgement w. congruency: Belief]
		- [Accounting for AI judgement w. congruency: Share]
		- [Optional w. congruency: Belief]
		- [Optional w. congruency: Share]

## Set all pairs of interest utilized throughout the code
```{r POI}
five_way_contrasts <- list(
	c("Forced_False_false", "Control_False_false"),
	c("Forced_False_unsure", "Control_False_unsure"),
	c("Forced_True_false", "Control_True_false"),
	c("Forced_True_unsure", "Control_True_unsure"),
	c("Forced_True_true", "Control_True_true")
)

optional_contrasts <- list(
  c("Opt_in_False_false", "Opt_out_False_false"),
  c("Opt_in_False_unsure", "Opt_out_False_unsure"),
  c("Opt_in_True_false", "Opt_out_True_false"),
  c("Opt_in_True_true", "Opt_out_True_true"),
  c("Opt_in_True_unsure", "Opt_out_True_unsure")
)
```

## Define function to print model stats
```{r p_func}
# Define function to extract overall p-value of model
model_p <- function(model_summary) {
    f <- model_summary$fstatistic
	p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL # Ensures the value is numeric
    if (p == 0) {
        p <- "< 2.2e-16"
    }
    return(p)
}

print_summary <- function(model) {
	model_summary <- summary(model)
	f_statistic <- model_summary$fstatistic[1]
	cat("F statistic:", f_statistic, "\n")
	
	df_residual <- model_summary$df[2]
	cat("Degrees of freedom (residuals):", df_residual, "\n")
	
	p_value <- model_p(model_summary)
	cat("P-value:", p_value, "\n")
	
	r2 <- model_summary$r.squared
	cat("R-squared:", r2, "\n")
	
	r2adj <- model_summary$adj.r.squared
	cat("Adjusted R-squared:", r2adj, "\n")
}

generate_contrast_vector <- function(pairs, levels) {
  # Initialize a zero vector of length equal to the number of levels
  contrast_vector <- rep(0, length(levels))
  
  # Find the indices of the pairs in the levels
  idx1 <- which(levels == pairs[1])
  idx2 <- which(levels == pairs[2])
  
  # Set the values in the contrast vector
  contrast_vector[idx1] <- 1
  contrast_vector[idx2] <- -1
  
  return(contrast_vector)
}
```

# Interactions

## Congruency

### Set equations (congruency)
```{r set_equations_int_congruency}
me_w_congruency_formula = exp_response ~ Condition*veracity*congruency + age + edu_recoded
five_way_congruency_formula = exp_response ~ Condition*cGPT_accuracy_recode*congruency + age + edu_recoded
optional_congruency_formula = exp_response ~ option_cond*cGPT_accuracy_recode*congruency + age + edu_recoded
```

### Main effects w. congruency: Belief
```{r me_congru_int_belief}
mb_me_congru <- lm(me_w_congruency_formula, data = belief_data)

print_summary(mb_me_congru)

clustered_se <- vcovCL(mb_me_congru, cluster = belief_data[, c("ResponseId", "qualtrics_question_num")])
round(coeftest(mb_me_congru, vcov = clustered_se),3)
```

### Main effects w. congruency: Share

```{r me_congru_int_share}
ms_me_congru <- lm(me_w_congruency_formula, data = share_data)

print_summary(ms_me_congru)

clustered_se <- vcovCL(ms_me_congru, cluster = share_data[, c("ResponseId", "qualtrics_question_num")])
round(coeftest(ms_me_congru, vcov = clustered_se),3)
```

### Accounting for AI judgement w. congruency: Belief

#### Fit regression model
```{r AI_judgement_congruency_belief}
mb_ai_congru <- lm(five_way_congruency_formula, data = belief_data_no_opt_or_hfc)

print_summary(mb_ai_congru)

clustered_se <- vcovCL(mb_ai_congru, cluster = belief_data_no_opt_or_hfc[, c("ResponseId", "qualtrics_question_num")])
round(coeftest(mb_ai_congru, vcov = clustered_se),3)
```

#### Post-hoc test

```{r AI_judgement_congruency_post_belief}
# Generate the estimated marginal means
test <- emtrends(mb_ai_congru, ~ Condition*cGPT_accuracy_recode, var="congruency", infer = TRUE)
test

# Extract the summary data frame from the emmGrid object
test_df <- summary(test)

# Create a vector of all combinations
temp_levels <- paste(test_df$Condition, test_df$cGPT_accuracy_recode, sep = "_")

# Create and name the contrast_list
contrast_list <- lapply(five_way_contrasts, generate_contrast_vector, levels = temp_levels)
names(contrast_list) <- sapply(five_way_contrasts, function(p) paste(p[1], "-", p[2]))

# Perform the specified pairwise comparisons using the generated contrast list
pairwise <- contrast(test, method = contrast_list, adjust = "bonferroni", infer = c(TRUE, TRUE))

# Convert the results to a data frame
ph_belief <- as.data.frame(pairwise)

# Round all numeric columns to three decimal places
ph_belief_rounded <- ph_belief
numeric_columns <- sapply(ph_belief, is.numeric)  # Identify numeric columns
ph_belief_rounded[numeric_columns] <- round(ph_belief[numeric_columns], 3)

# Print the rounded results
print(ph_belief_rounded)

plot_model(mb_ai_congru, type  = "pred", terms =  c( "congruency", "Condition",  "cGPT_accuracy_recode"), title = "Moderting effects of headline congurency") + ylab("Intention to share") + theme_minimal()
```

### Accounting for AI judgement w. congruency: Share

#### Fit regression model
```{r AI_judgement_congruency_share}
ms_ai_congru <- lm(five_way_congruency_formula, data = share_data_no_opt_or_hfc)

print_summary(ms_ai_congru)

clustered_se <- vcovCL(ms_ai_congru, cluster = share_data_no_opt_or_hfc[, c("ResponseId", "qualtrics_question_num")])
round(coeftest(ms_ai_congru, vcov = clustered_se),3)
```

#### Post-hoc test
```{r AI_judgement_congruency_post_share}
# Generate the estimated marginal means
test <- emtrends(ms_ai_congru, ~ Condition*cGPT_accuracy_recode, var="congruency", infer =TRUE)
test

# Extract the summary data frame from the emmGrid object
test_df <- summary(test)

# Create a vector of all combinations
temp_levels <- paste(test_df$Condition, test_df$cGPT_accuracy_recode, sep = "_")

# Create and name the contrast_list
contrast_list <- lapply(five_way_contrasts, generate_contrast_vector, levels = temp_levels)
names(contrast_list) <- sapply(five_way_contrasts, function(p) paste(p[1], "-", p[2]))

# Perform the specified pairwise comparisons using the generated contrast list
pairwise <- contrast(test, method = contrast_list, adjust = "bonferroni", infer = c(TRUE, TRUE))

# Convert the results to a data frame
ph_share <- as.data.frame(pairwise)

# Round all numeric columns to three decimal places
ph_share_rounded <- ph_share
numeric_columns <- sapply(ph_share, is.numeric)  # Identify numeric columns
ph_share_rounded[numeric_columns] <- round(ph_share[numeric_columns], 3)

# Print the rounded results
print(ph_share_rounded)

plot_model(ms_ai_congru, type  = "pred", terms =  c( "congruency", "Condition",  "cGPT_accuracy_recode"), title = "Moderting effects of headline congurency") + ylab("Intention to share") + theme_minimal()
```

### Optional w. congruency: Belief

#### Fit regression model
```{r optional_congruency_belief}
mb_opt_congru <- lm(optional_congruency_formula, data = belief_data_opt)

print_summary(mb_opt_congru)

clustered_se <- vcovCL(mb_opt_congru, cluster = belief_data_opt[, c("ResponseId", "qualtrics_question_num")]) 
round(coeftest(mb_opt_congru, vcov = clustered_se),3)
```

#### Post-hoc test
```{r optional_congruency_post_belief}
# Generate the estimated marginal means
test <- emtrends(mb_opt_congru, ~ option_cond* cGPT_accuracy_recode, var="congruency", infer= TRUE)
test

# Extract the summary data frame from the emmGrid object
test_df <- summary(test)

# Create a vector of all combinations
temp_levels <- paste(test_df$option_cond, test_df$cGPT_accuracy_recode, sep = "_")

# Create and name the contrast_list
contrast_list <- lapply(optional_contrasts, generate_contrast_vector, levels = temp_levels)
names(contrast_list) <- sapply(optional_contrasts, function(p) paste(p[1], "-", p[2]))

# Perform the specified pairwise comparisons using the generated contrast list
pairwise <- contrast(test, method = contrast_list, adjust = "bonferroni", infer = c(TRUE, TRUE))

# Convert the results to a data frame
ph_belief <- as.data.frame(pairwise)

# Round all numeric columns to three decimal places
ph_belief_rounded <- ph_belief
numeric_columns <- sapply(ph_belief, is.numeric)  # Identify numeric columns
ph_belief_rounded[numeric_columns] <- round(ph_belief[numeric_columns], 3)

# Print the rounded results
print(ph_belief_rounded)

plot_model(mb_opt_congru, type  = "pred", terms =  c( "congruency", "option_cond",  "cGPT_accuracy_recode"), title = "Moderting effects of headline congurency") + ylab("Believed") + theme_minimal()
```

### Optional w. congruency: Share

#### Fit regression model
```{r optional_congruency_share}
ms_opt_congru <- lm(optional_congruency_formula, data = share_data_opt )

print_summary(ms_opt_congru)

clustered_se <- vcovCL(ms_opt_congru, cluster = share_data_opt[, c("ResponseId", "qualtrics_question_num")]) 
round(coeftest(ms_opt_congru, vcov = clustered_se),3)
```

#### Post-hoc test
```{r optional_congruency_post_share}
# Generate the estimated marginal means
test <- emtrends(ms_opt_congru, ~ option_cond*cGPT_accuracy_recode, var="congruency", infer = TRUE)
test

# Extract the summary data frame from the emmGrid object
test_df <- summary(test)

# Create a vector of all combinations
temp_levels <- paste(test_df$option_cond, test_df$cGPT_accuracy_recode, sep = "_")

# Create and name the contrast_list
contrast_list <- lapply(optional_contrasts, generate_contrast_vector, levels = temp_levels)
names(contrast_list) <- sapply(optional_contrasts, function(p) paste(p[1], "-", p[2]))

# Perform the specified pairwise comparisons using the generated contrast list
pairwise <- contrast(test, method = contrast_list, adjust = "bonferroni", infer = c(TRUE, TRUE))

# Convert the results to a data frame
ph_belief <- as.data.frame(pairwise)

# Round all numeric columns to three decimal places
ph_belief_rounded <- ph_belief
numeric_columns <- sapply(ph_belief, is.numeric)  # Identify numeric columns
ph_belief_rounded[numeric_columns] <- round(ph_belief[numeric_columns], 3)

# Print the rounded results
print(ph_belief_rounded)

plot_model(ms_opt_congru, type  = "pred", terms =  c( "congruency", "option_cond",  "cGPT_accuracy_recode"), title = "Moderting effects of headline congurency") + ylab("Intention to share") + theme_minimal()
```
